Hydra Battles and Cie

### 2021-05-14 doc/chapter-primrec.tex
proof that the length  of hydra battles is not primitive recursive

>>>>

\subsection{The length of standard hydra battles}

The module \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra\_Theorems} contains a proof that the function which computes the lenght of standard hydra battles is not primitive recursive. This proof is  a little more complex than the preceding one.

\subsubsection{Definitions}

The function we consider is defined and proven correct in
Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length}.

\begin{Coqsrc}
Definition l_std alpha k := (L_ alpha (S k) - k)%nat.

Lemma l_std_ok : forall alpha : E0,
    alpha <> Zero ->
    forall k : nat,
      1 <= k -> battle_length standard k (iota (cnf alpha))
                              (l_std alpha k).
(* ... *)
\end{Coqsrc}

\subsubsection{Proof steps}

Now, let us assume that the function \texttt{l\_std} is primitive recursive.

\emph{From \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}.

\begin{Coqsrc}
Section battle_lenght_notPR.

  (** We assume that the function with computes the length 
      of standard battles is primitive recursive *)

  Hypothesis H: forall alpha, isPR 1 (l_std alpha). 
\end{Coqsrc}

Let us consider the hydra represented by the ordinal $\omega^\omega$.

\begin{Coqsrc}
Let alpha := Phi0 omega%e0.
Let h := iota (cnf alpha).
\end{Coqsrc}

In order to get rid of the substraction in the definition of \texttt{l\_std}, we work with a helper function.

\begin{Coqsrc}
Let m k := L_ alpha (S k).

Remark m_eqn : forall k, m k = (l_std alpha k + k)%nat.
(* ... *)
\end{Coqsrc}

Under the hypothesis \texttt{H}, $m$ is also primitive recursive.

\begin{Coqsrc}
Remark mIsPR : isPR 1 m.
(* ... *)
\end{Coqsrc}

Now, let us recall the following lemmas :

\begin{enumerate}
\item 
\emph{From \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}}

\begin{Coqsrc}
H_F : forall (alpha : E0) (n : nat), 
     F_ alpha (S n) <= H_ (Phi0 alpha) (S n) 
\end{Coqsrc}

\emph{Please note that our proof of this lemma is not trivial at all. It uses some lemmas of Chapters~\vref{chap:ketonen} and \vref{chap:alpha-large}.}

\item
\noindent
\emph{From \href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}

\begin{Coqsrc}
H_L_ : forall (alpha : E0) (i : nat), H_ alpha i <= L_ alpha (S i)
\end{Coqsrc}

\item

\emph{From \href{../theories/html/hydras.Epsilon0.F_omega.html}{Epsilon0.F\_omega}}

\begin{Coqsrc}
F_vs_Ack : forall n : nat, 2 <= n -> Ack n n <= F_ omega n
\end{Coqsrc}

\end{enumerate}

By transitivity, we get the inequality
$F_\omega(k+1)\leq m(k+1)$, for any $k$.

\begin{Coqsrc}
Remark m_ge_F_omega : forall k,  F_ omega (S k) <= m (S k).
\end{Coqsrc}

We finish the proof by noting that the function $m$ (composed with \texttt{S}) dominates the Ackermann function, which leads to a contradiction.

\begin{Coqsrc}
 Remark m_dominates_Ack : 
     dominates (fun n =>  S (m n)) (fun n => Ack.Ack n n).
(* ... *)

 Remark SmNotPR : isPR 1 (fun n => S (m n)) -> False.

 Theorem LNotPR : False.
  Proof.
    apply SmNotPR,  compose1_1IsPR.
    - apply mIsPR.
    - apply succIsPR.
  Qed.

End battle_lenght_notPR.
\end{Coqsrc}
